### 2019-10-04 s1_text/s1_text.tex

\documentclass[10pt,letterpaper]{article}
\usepackage[top=1.2in,left=1.65in,right=1.65in,footskip=0.6in]{geometry}

% amsmath and amssymb packages, useful for mathematical formulas and symbols
\usepackage{amsmath,amssymb}

% Use adjustwidth environment to exceed column width (see example table in text)
\usepackage{changepage}

% Use Unicode characters when possible
\usepackage[utf8x]{inputenc}

% textcomp package and marvosym package for additional characters
\usepackage{textcomp,marvosym}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

% Use nameref to cite supporting information files (see Supporting Information section for more info)
\usepackage{nameref,hyperref}

% line numbers
\usepackage[right]{lineno}

% ligatures disabled
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% color can be used to apply background shading to table cells only
\usepackage[table]{xcolor}

% array package and thick rules for tables
\usepackage{array}

% create "+" rule type for thick vertical lines
\newcolumntype{+}{!{\vrule width 2pt}}

% create \thickcline for thick horizontal lines of variable length
\newlength\savedwidth
\newcommand\thickcline[1]{%
	\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
	\cline{#1}%
	\noalign{\vskip\arrayrulewidth}%
	\noalign{\global\arrayrulewidth\savedwidth}%
}

% \thickhline command for thick horizontal lines that span the table
\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
	\hline
	\noalign{\global\arrayrulewidth\savedwidth}}

% Remove comment for double spacing
%\usepackage{setspace} 
%\doublespacing

% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
% Captions will be left justified
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
\renewcommand{\figurename}{Fig}

% Use the PLoS provided BiBTeX style
\bibliographystyle{plos2015}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Figures
\usepackage{graphicx}

%% Include all macros below

\usepackage{placeins}
\usepackage{refcount}
\usepackage{textgreek}
\usepackage{bm}
\usepackage{url} % url in Quellen verlinken

\renewcommand\thesection{\Alph{section}}

\newcommand{\unit}[1]{\,\mathrm{#1}}
\newcommand{\n}[1]{\mathrm{#1}}
\newcommand{\dd}[2]{\frac{\mathrm{d} #1}{\mathrm{d} #2}}
\newcommand*{\defeq}{\mathrel{\vcenter{\baselineskip0.5ex \lineskiplimit0pt
			\hbox{\scriptsize.}\hbox{\scriptsize.}}}%
	=}
\newcommand\subref[2]{%
	\def\myref{\getrefnumber{#1}}% extract the reference number
	\hyperref[#1]{\myref\mbox{#2}}% print the reference number with argument
}

\hyphenation{XapA}
\hyphenation{XapB}
\hyphenation{XapR}
\hyphenation{XapAB}
\hyphenation{XapABR}
\hyphenation{xapA}
\hyphenation{xapB}
\hyphenation{xapR}
\hyphenation{xapAB}
\hyphenation{xapABR}

\usepackage[obeyFinal]{todonotes} % todo notes and missing figures

%% END MACROS SECTION

\begin{document}

\begin{flushleft}
	{\Large
		\textbf\newline{Supporting information for ``theoretical investigation of a genetic switch for metabolic adaptation''}
	}
	\newline
	% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
	\\
	Kathrin Laxhauber\textsuperscript{1},
	Muir J. Morrison\textsuperscript{2},
	Rob Phillips\textsuperscript{2,3$\ast$}
	\\
	\bigskip
	\textbf{1} Dept???, ETHZ, Zurich, State???, Switzerland\todo{MJM: Add dept \& complete affiliation}
	\\
	\textbf{2} Department of Physics, California Institute of Technology, Pasadena, CA, USA
	\\
	\textbf{3} Department of Biology and Biological Engineering, California Institute of Technology, Pasadena, CA, USA
	\\
	\bigskip

	% Insert additional author notes using the symbols described below. Insert symbol callouts after author names as necessary.
	% 
	% Remove or comment out the author notes below if they aren't used.
	%
	% Primary Equal Contribution Note
	% Yinyang These authors contributed equally to this work.

	% Additional Equal Contribution Note
	% Also use this double-dagger symbol for special authorship notes, such as senior authorship.
	% \ddag These authors also contributed equally to this work.

	% Current address notes
	% \textcurrency Current Address: Dept/Program/Center, Institution Name, City, State, Country % change symbol to "\textcurrency a" if more than one current address note
	% \textcurrency b Insert second current address 
	% \textcurrency c Insert third current address

	% Use the asterisk to denote corresponding authorship and provide email address in note below.
	$\ast$ phillips@pboc.caltech.edu

\end{flushleft}

\newpage
\linenumbers

\section{Model choices, simplifications and assumptions}
This part of supporting information text offers a more detailed discussion
of the assumptions that were made in the model and explains the choices that
were made.

\paragraph*{Simplistic modeling of transcription.}
The precise mechanism at the promoter level is not the key feature of this
system, and we expect the same qualitative results for slightly different
transcription models. This work aims to gain an intuition from the results,
and the more unknown parameters there are, the harder this becomes.
Furthermore, the mechanistic details of transcription are not entirely known
and would make the model too complex: for example, transcription probably is
not a Poisson process, but comes in bursts. Additionally, very little is
known about the function of the \emph{xapAB} promoter. For these reasons, we
do not model every detail, but focus on the most relevant features of the
system, while still trying to be physically correct.

Firstly, the equilibrium assumption is not completely accurate, but has
proven successful in previous work. An example for this is the very detailed
analysis of a gene with one repressor in~\cite{Phillips2018}. Secondly,
different binding energies of XapR to its two binding sites would change the
equations as follows: the factor $\left(\frac{1}{K_{\n{XapR}}}\right)^2$
would become $\frac{1}{K_{\n{XapR,1}} K_{\n{XapR,2}}}$ and the factor
$\frac{2}{K_{\n{XapR}}}$ would be replaced by $\frac{K_{\n{XapR,1}} +
K_{\n{XapR,2}}}{K_{\n{XapR,1}} K_{\n{XapR,2}}}$. For most choices of
$K_{\n{XapR}}$, this does not give the exact same results quantitatively,
but we do observe the same qualitative outcome (see also next paragraph).
Thirdly, we consider all but one state as inactive, which is not strictly
correct. Still, as mentioned in the main text, experiments suggest that this
is reasonable. The fourth main simplification, which is the independence of
polymerase and XapR binding, will be discussed in a separate subsection
below.

In Fig~\ref{figS1:transcription}, we have plotted phase portraits where some
of the above assumptions have not been made. For that, we have chosen some
extreme values for the new parameters to demonstrate that these
simplifications do not have a significant influence. The two additional
transcription rates are large compared to experimental observations.
Moreover, the chosen values of $\n{[XapR]}_{\n{R},i}$ correspond to a
difference of two orders of magnitude between $K_{\n{XapR,1}}$ and
$K_{\n{XapR,2}}$. The difference between Oid and O3 in \emph{lac} are two to
three orders of magnitude~\cite{RazoMejia2018}. The values were chosen such
that $K_{\n{XapR,1}} K_{\n{XapR,2}} = K_{\n{XapR}}^2$. Unsurprisingly, this
is necessary to not observe large deviations from the original equations.
The results lead to the conclusion that, for the purposes of this paper, the
aforementioned simplifications are appropriate.

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{FigSI1.pdf}
	\caption{{\bf Results for different models of transcription.}
		The parameters are the same as the ones that were used for all other
		phase portraits in this report and that are listed in a table in the
		main part. In (A), the phase portrait after addition of non-zero
		transcription rates to all other polymerase-bound states is shown
		(third assumption lifted). The new parameter values are $r_{\n{I}} =
		10^{-2} r_{\n{p}}$ for the transcription rate when only one XapR is
		bound and $r_{\n{0}} = 10^{-3} r_{\n{p}}$ for the basal rate. In
		(B), the phase portrait after introducing different XapR
		dissociation constants can be seen (second assumption lifted). The
		new parameter values are $\n{[XapR]_{R,1}} = 10$ and
		$\n{[XapR]_{R,2}} = 0.1$.}
	\label{figS1:transcription}
\end{figure}

\paragraph*{Activator and polymerase interaction.}
A mechanistically more precise model of activation would include an
interaction energy between polymerase and XapR. In that case,
$p_{\n{active}}$ cannot be factored into separate terms for the polymerase
and the XapR binding probability. However, the precise activation mechanism
is unknown.

Moreover, introducing an interaction energy between polymerase and XapR
would make the equations overly complex. This would not only make
factorization impossible, it would also be just as consequential to have
different dissociation constants and additional rates as in the previous
paragraph. This leads to a huge number of terms and parameters because of
different interaction energies for every state, which can be foreseen from
the states and weights shown in the main text. The precise transcription
mechanism is not one of the central parts of the circuit we are
investigating (see also discussion in the previous paragraph). Thus, making
it the by far most complex and precise part of the equations would be
unreasonable and hinder the understanding of the more relevant components of
the system.

\begin{table}
	\centering
	\caption{
		{\bf States and weights for a more precise model of transcription.}}
	\begin{tabular}{llll}
		\thickhline
		XapR & Pol. & Transcription & Weight \\
		\hline
		none & no & $0$ & $1$ \\
		first & no & $0$ & $w_{\n{X1}}$ \\
		second & no & $0$ & $w_{\n{X2}}$ \\
		both & no & $0$ & $w_{\n{X1}} w_{\n{X2}} e^{\Delta \epsilon_{\n{coop}}}$ \\
		none & yes & $r_0$ & $w_{\n{p}}$ \\
		first & yes & $r_{\n{X1}}$ & $w_{\n{p}} w_{\n{X1}} e^{\Delta \epsilon_{\n{X1}}}$ \\
		second & yes & $r_{\n{X2}}$ & $w_{\n{p}} w_{\n{X2}} e^{\Delta \epsilon_{\n{X2}}}$ \\
		both & yes & $r_2$ & $w_{\n{p}} w_{\n{X1}} w_{\n{X2}} e^{\Delta \epsilon_{\n{X1}} + \Delta \epsilon_{\n{X2}} + \Delta \epsilon_{\n{coop}}}$ \\
		\thickhline
	\end{tabular}
	\begin{flushleft} 
		The first column shows which XapR sites are occupied in this state,
		the second one whether polymerase is bound to the promoter, and the
		third the transcription rate. In the weights, the notation is as
		follows: $w_{\n{X1}}$ is the weight for XapR binding to the first
		binding site on the promoter, $w_{\n{X2}}$ is that to the second
		site, $w_{\n{p}}$ is the weight of polymerase binding to the
		promoter, $e^{\Delta \epsilon_{\n{X1}}}$ is the interaction energy
		between one XapR on the first binding site and the polymerase that
		is bound, $e^{\Delta \epsilon_{\n{X2}}}$ is the same for the second
		binding site, and finally, $e^{\Delta \epsilon_{\n{coop}}}$ is the
		interaction energy of the two XapR at both binding sites as in the
		main part.
	\end{flushleft}
	\label{tableSI1:statesPromoter}
\end{table}

Furthermore, we do not expect our approximation to have a large influence:
the interaction energy will make the doubly bound state dominate, and the
polymerase-XapR interaction should be weak in the singly bound state,
because very weak activation is observed in the experiment when one of the
XapR binding sites is removed~\cite{Chure2019}. If $e^{\Delta
\epsilon_{\n{X1}}}$ and $e^{\Delta \epsilon_{\n{X2}}}$ are small, we have
$w_{\n{p}} w_{\n{X1}} e^{\Delta \epsilon_{\n{X1}}} \approx w_{\n{p}}
w_{\n{X2}} e^{\Delta \epsilon_{\n{X2}}} \approx w_{\n{p}} w_{\n{X}}$. If, in
addition, the transcription rates for the singly bound states are much
smaller than those for the doubly bound state, we can neglect the singly
bound states in the nominator of $p_{\n{active}}$ (which, as argued in the
previous paragraph ``Simplistic modeling of transcription'', is a good
assumption). Thereby, we can obtain the same form of $p_{\n{active}}$ that
we are working with in the main part by including the interaction between
polymerase and XapR into $\epsilon_{\n{coop}}$.

\paragraph*{Working with only one protein species.}
The proteins are transcribed together, and thus, their rate of transcription
and mRNA decay should be the same. The protein decay is assumed to be
dominated by cell division, making the decay rates roughly equal as well.
What could differ is the translation rate. Changes in this rate rescale the
protein synthesis rate by the same factor and thus, in steady-state, change
the total number of proteins. The only other place where the proteins appear
in the ODE's is in their respective Michaelis-Menten term, and the latter is
proportional to $k_{\n{cat}} \n{[E]_0}$. Hence, different numbers of XapA
and XapB can be accounted for by changing their respective values of
$k_{\n{cat}}$.

\paragraph*{Exponential protein decay.}
The protein decay should be dominated by dilution through cell division,
which is a discrete process. However, the cell grows before division,
resulting in a continuous dilution. Furthermore, replication leads to a
varying number of gene copies and thereby also a varying transcription rate.
Trying to capture all these mechanistic details would lead to an
overcomplicated model. Treating the protein decay as a Poisson process has
produced correct results in past research, which is why we expect this
assumption to be reasonable.

\paragraph*{Kinetics of XapB, NupC, and NupG.}
All transporters in our system are driven by the proton
gradient~\cite{Norholm2001}, and we assume their kinetic scheme to be like
that of \emph{lac} permease. This scheme is shown schematically in
Fig~\ref{figS2:permease}. Even though it includes several steps, the
essential part is substrate binding, conformational change of the
transporter, and substrate release on the other side of the membrane. We
assume the proton gradient across the membrane to be constant and proton
binding and unbinding to the transporter to be fast. Hence, the latter does
not need to be included explicitly in the kinetic model (it is implicitly
part of the turnover rate).

\begin{figure}
	\centering
	\includegraphics[width=0.7\textwidth]{FigSI2.pdf}
	\caption{{\bf Kinetic scheme for transport by proton-driven pumps.}
		The mechanism of lac permease as for example described
		in~\cite{Kaback2015} is used as a model for xanthosine transport.
		The enzyme binds a proton and a substrate (steps 1 and 2), undergoes
		conformational change (steps 3 and 4), and then releases substrate
		and proton on the other side of the membrane (steps 5 and 6). When
		the enzyme is empty, it can change its conformation again to repeat
		the transport (steps 7 and 8). Instead, it could also bind a new
		proton and substrate and transport them in the opposite direction
		(steps 6, 5, 4, 3, 2, 1). Alternatively, it could also keep the
		proton and just cycle back and forth, exchanging substrates from
		outside and inside the cell (steps 5, 4, 3, 2).}
	\label{figS2:permease}
\end{figure}

Substrate release should be much faster than conformational change, which is
analogous to ignoring the ``EP'' state in Michaelis-Menten (enzyme with
product bound, here the conformationally changed transporter with substrate
still bound). Also, it should be reasonable to neglect the immediate
backwards reaction (before proton release), and thus, Michaelis-Menten
kinetics is obtained. Data from~\cite{Kaczorowski1979}
and~\cite{Viitanen1984} actually show that, at a physiologically reasonable
proton gradient, both, influx and efflux, can be described by the
Michaelis-Menten equation very well.

What is neglected in this modeling is the fact that the transporter needs to
release the proton and undergo a conformational change again before being
able to conduct the same reaction again. However, by using experimentally
determined apparent values to estimate the kinetic constants, this is
already included in these parameter values. Another simplification we make
is to treat influx and efflux as two independent reactions instead of
thinking of it as one reversible kinetic process. This seems reasonable,
though, because the transporter needs to lose its proton between reactions
in order to obtain a net transport. If the transporter has for example done
steps 1-6 in Fig~\ref{figS2:permease}, it can from there on really enter two
entirely different reactions: it can do its conformational change in its
empty state (steps 7 and 8) to then transport a substrate again (steps 1-6)
or it can bind a new proton and a new substrate directly and transport them
back (steps 6-1). The former would give influx and the latter efflux, so the
two of them can be modeled as separate reactions.

The NupC and NupG reactions are approximated as linearly dependent on the
substrate. This should be appropriate, since their affinity for xanthosine
is so low that it cannot even be measured properly~\cite{Norholm2001}. This
makes $K_{\n{M}}$ much larger than $c$ or $\n{[x]}$ will ever be. For XapB,
this is not the case, because it has much higher specificity for xanthosine
and $c$ and $\n{[x]}$ take values around the $K_{\n{M}}$ values of XapB.

\paragraph*{Passive transport across the membrane.}  
Most small molecules diffuse passively through the membrane, but at
physiological pH, xanthosine is mostly negatively charged and charged
substances do not significantly diffuse across the membrane on their own.
Nevertheless, from the $\n{pK_a}$ of xanthosine we find that at
physiological pH, about 1.5\% of it is uncharged and could pass the
membrane. This could be modeled as follows:
\begin{eqnarray}
\label{eq:diff}
\left(\dd{\n{[x]}}{t}\right)_{\n{diff}} = J A = p (c-\n{[x]}) A = \underbrace{p A}_{\n{rate}} (c-\n{[x]})
\end{eqnarray}
Here, $J$ is the flux through the membrane, which follows directly from
Fick's first law (if steady-state diffusion is assumed). The other new
parameters are the membrane area $A$ and the membrane permeability
coefficient $p$ of the diffusing substrate.

The diffusion rate $k_{\n{d}} = p A$ can be estimated with the help of
Overton's rule. In our model, we find that this passive diffusion alone
(i.e. without NupC and NupG) is enough to obtain bistability in the
deterministic analysis as well as bimodality in the stochastic simulations.
However, experiments show that \textDelta\emph{nupC} \textDelta\emph{nupG}
strains cannot grow on xanthosine~\cite{Norholm2001}. This suggests that
passive diffusion alone is not enough to make the cells switch. Experiments
with cells in minimal media containing some glucose and a xanthosine
concentration in the bimodal regime show that the switching time can be
several generations. The same is observed in our simulations. The xanthosine
concentration that was used for the double knockout experiments, where no
glucose was present, is just above the switching threshold, which is when
switching is slow. For this reason, we suspect that without NupC and NupG
and with only xanthosine as an energy resource, the cells starve before they
have switched. This probably is the reason why passive diffusion is, in
reality, not enough to observe switching. NupC and NupG can probably keep
the cells going at very low intracellular xanthosine concentrations. Because
of the active transport, they can accumulate a higher intracellular than
extracellular concentration of xanthosine.

Since NupC and NupG transport seems to be much more significant than
diffusion, we do not include the latter explicitly in our model. However,
because it takes the form given in Eq~\ref{eq:diff}, it can easily be
accounted for implicitly by slightly changing the values of $k_{\n{nup}}$
and $\xi$.

\paragraph*{Other mechanisms for xanthosine degradation.}
There probably are some ways in which xanthosine is degraded, other than by
XapA. Otherwise, initially uninduced cells that are transferred to only
xanthosine as an energy source might not survive until they have switched.
As mentioned previously, it was found that they do~\cite{Norholm2001}, so
they need to be able to metabolize xanthosine somehow. Nevertheless, these
degradation pathways cannot have a big effect: as argued later on when
estimating the Nup parameters, the intracellular xanthosine concentration in
the uninduced state must be a bit larger, but not too much larger, than the
extracellular one. Thus, degradation of xanthosine has to be small because
import through Nup is weak. Due to this, it should be appropriate to
linearize the equations describing these effects. Then, they can simply be
accounted for in the term for Nup transport by making slight changes in the
values of $\xi$ and $k_{\n{nup}}$. It could, however, also be that the basal
level of XapA is enough to obtain the small degradation effect without any
other pathways.

\section{Parameter estimation}
In this section we present a detailed estimation of the values of all
nondimensional parameters. Two dimensionful parameters appear in most
nondimensional ones, namely $\gamma_{\n{p}}$ and $K_{\n{a}}$. For this
reason, we start by estimating these.

\paragraph*{The protein decay rate $\bm{\gamma_{\n{p}}}$:}
Protein decay is generally dominated by dilution through cell growth and
division. The average doubling time of \emph{E. coli} is 20-30 minutes. We
take this as an estimate for the protein half life, which leads us to a
decay rate of $\gamma_{\n{p}} = \frac{\ln 2}{t_{1/2}} \approx 5\substack{+2
\\ -3} \cdot 10^{-4}~\n{s^{-1}}$. Note that these upper and lower bounds
correspond to half-lifes between 15 and 60 minutes, which should be
reasonable. For the following nondimensionalization, we work with a fixed
value of $\gamma_{\n{p}} = 5 \cdot 10^{-4}~\n{s^{-1}}$.

\paragraph*{The Michaelis-Menten parameters for XapA ($K_{\n{a}}$, $k_{\n{\alpha}}$):}
The Michaelis-Menten parameters of XapA kinetics have been measured in two
independent publications~\cite{Dandanell2005,Koszalka1988}. Both have found
similar values and we conclude $K_{\n{a}} \approx 6 \pm 3 \cdot
10^{-5}~\n{M}$ and $k_{\n{a}} \approx 10^{-1 \pm 0.8}~\n{s^{-1}}$. This
gives $k_{\n{\alpha}} \approx 10^{2 \pm 0.8}$. We choose to work with
$K_{\n{a}} = 5 \cdot 10^{-5}~\n{M}$ in the following.

\paragraph*{The extracellular xanthosine concentration $\n{[c]_{a}}$:} In
the experiment, switching occurred if cells were grown in a solution with a
concentration of xanthosine of roughly a few mM~\cite{Norholm2001}.
According to these experimental observations, the interesting range of
xanthosine should be $\n{[c]_{a}} \in [ 0, 10^3 ]$ (nondimensionalized),
which corresponds to $\n{c} \in [ 0, 5 \cdot 10^{-2} ]~\n{M}$. However, we
are careful because there are issues with the solubility of xanthosine.

\paragraph*{The MWC parameters $K_{\n{\chi A}}$, $K_{\n{IA}}$, and $\Delta \epsilon_{\n{x}}$:}
The three MWC parameters are the energy barrier between the active and
inactive state of XapR and the equilibrium constants of xanthosine binding
to XapR in the two states. In Fig~\ref{figS2:induction}, the induction curve
as a function of the nondimensional parameters is shown. We assume that
evolution has tuned the induction curve such that less than $10 \%$ of XapR
is active in the absence of xanthosine, and more than $90 \%$ is active for
$\n{[x]}$ approaching infinity. This is a relatively weak assumption,
otherwise XapR would not really be fulfilling its job. It helps us by
putting a constraint on $K_{\n{IA}}$ and $\Delta \epsilon_{\n{x}}$:
\begin{alignat*}{1}
p_{\n{active}}(\n{[x]_a} = 0) &= \frac{1}{1 + e^{\Delta \epsilon_{\n{x}}}} < 0.1 \\
1-p_{\n{active}}(\n{[x]_a} \to \infty) &= \frac{e^{\Delta \epsilon_{\n{x}}}}{K_{\n{IA}}^2 + e^{\Delta \epsilon_{\n{x}}}} < 0.1
\end{alignat*}
Thus, 
\begin{eqnarray*}
	\Delta \epsilon_{\n{x}} > 2.2 \hspace{2 em} \n{and} \hspace{2 em} K_{\n{IA}} > 3 e^{\Delta \epsilon_{\n{x}} / 2}.
\end{eqnarray*}

A further constraint is found from $K_{\n{IA}}=\frac{K_{\n{x I}}}{K_{\n{x
A}}} = e^{\beta \Delta E}$, where $\Delta E$ is the energy difference
between xanthosine binding to the two states. It is unlikely that this
energy difference is more than $7~k_{\n{B}} T$, which corresponds to roughly
three strong hydrogen bonds. This gives us an upper bound $K_{\n{IA}} <
10^3$ and thereby $\Delta \epsilon_{\n{x}} < 10^1$. For $K_{\n{IA}} = 10^2$,
the upper bound is $\Delta \epsilon_{\n{x}} < 7$. This gives us a range of
$K_{\n{IA}} \approx 10^1 - 10^3$ and $\Delta \epsilon_{\n{x}} \approx 2 -
10$ or, more precisely, $\Delta \epsilon_{\n{x}} \approx 2$ to
$\ln\left(\frac{1}{9} \left(K_{\n{IA}}\right)^2 \right) \approx 2 \left(
\ln\left(K_{\n{IA}}\right) - 1\right) < 12$. Compared with other systems,
these values do seem reasonable.

\begin{figure}
	\centering
	\includegraphics[width=0.85\textwidth]{FigSI3.pdf}
	\caption{{\bf Induction curves given by MWC model.}
	The main plot shows the probability of XapR being in the active state as
	a function of the logarithm of the xanthosine concentration for $\Delta
	\epsilon_{\n{x}}$ as shown, $K_{\n{\chi A}} = 10^{2}$, and $K_{\n{IA}} =
	10^{2}$. The shaded areas indicate where the curve lies for changes in
	$K_{\n{IA}}$ of one order of magnitude. Changing $K_{\n{\chi A}}$ by one
	order of magnitude, on the other hand, shifts all the curves by one
	order of magnitude in the corresponding direction to the left or right.
	The inset shows how the allowed regimes of $\Delta \epsilon_{\n{x}}$ and
	$K_{\n{IA}}$ depend on each other. This explains why for $\Delta
	\epsilon_{\n{x}} = 7$ and $K_{\n{IA}}=10$ (lower green line in the main
	plot), $p_{\n{active}}$ at infinite xanthosine concentrations becomes
	far too low.}
	\label{figS2:induction}
\end{figure}

To find $K_{\n{\chi A}}$ we furthermore assume that the steep part of the
induction curve lies in the biologically relevant xanthosine regime, which
estimated experimentally, see the paragraph about $\n{[c]_a}$ above. Again,
this is a rather weak assumption. As can be seen from
Fig~\ref{figS2:induction}, this assumption sets the value range of
$K_{\n{\chi A}}$. The line for $\Delta \epsilon_{\n{x}} = 5$ in
Fig~\ref{figS2:induction} is at the experimentally expected position in the
xanthosine regime. It becomes clear from the figure that the position of the
steep range is strongly affected by $\Delta \epsilon_{\n{x}}$, and a change
of 1 in the latter leads to roughly a change of $10^1$ in the former. Thus,
we estimate $K_{\n{\chi A}} \approx 10^{2 \pm 1} \cdot 10^{\Delta
\epsilon_{\n{x}}-5}$ and thereby finish our estimation of the MWC
parameters.

\paragraph*{The XapR concentration $\n{[XapR]_R}$:} 
It is experimentally known that the XapR copy number is very low, on the
order of $10^{-8}~\n{M}$ ($10^1$ molecules per cell)~\cite{Li2014}. We
estimate the XapR -- DNA dissociation constant from the values for
\emph{lac} that were found in~\cite{RazoMejia2018} and obtain $10^{-8 \pm
2}~\n{M}$, where we have added one additional order of magnitude as
insecurity to the upper and lower bound. This gives
$\n{[XapR]_R}=\frac{\n{[XapR]_{\n{tot}}}}{K_{\n{XapR}}} \approx 10^{0 \pm
2}$.

\paragraph*{The XapR cooperative energy $\Delta \epsilon_{\n{coop}}$:}
Any cooperative energy should come mainly from an interaction between the
two XapR molecules. It would be surprising if it was larger than $10 \
k_{\n{B}} T$, which corresponds to roughly 4 strong hydrogen bonds. Hence,
we estimate $\Delta \epsilon_{\n{coop}} \approx 0 - 10$.

\paragraph*{The transcription rate parameter $\rho_{\n{m}}$:} The
transcription rate $r_{\n{m}} \frac{\n{[P]}}{\n{[P]} + K_{\n{P}}}$ in the
presence of two activator molecules should be on the order of $10^{-2 \pm
1}~\n{nM\ s^{-1}}$, thus $\rho_{\n{m}} \defeq
\frac{r_{\n{m}}}{\gamma_{\n{p}} K_{\n{a}}} \frac{\n{[P]}}{K_{\n{P}}+\n{[P]}}
\approx 10^{-3 \pm 2}$. Note that the units are $\n{nM\ s^{-1}}$ and not
just $\n{s^{-1}}$ because the corresponding rate equation is of the form
$\dd{\n{[m]}}{t} = r \cdot p$ where the probability $p$ is dimensionless and
the left-hand side has dimensions of mRNA concentration per time.

This estimate comes from the assumption that this rate should be roughly
$10^{1} - 10^{2}$ times larger than a normal basal rate (of other genes),
for which common values are around $10^{-3 \pm 1}~\n{nM\
s^{-1}}$~\cite{Milo2016}. Because experiments show that compared to the
activated promoter, the basal transcription rate is almost zero for
\emph{xapAB} (probably due to the weak polymerase binding site), this
actually means a $10^{2}$- to $10^{3}$-fold increase in the transcription
rate due to activation by XapR.

Keep in mind that the rate includes not only the effect of XapR on the pure
speed of transcription, bt also any effect of XapR on the polymerase
dissociation constant. Still, we do not expect higher values such as
$10^{0}~\n{nM\ s^{-1}}$ since this is roughly the rate for transcription of
rRNA and the latter should be significantly faster. To further validate our
estimate, we can compare it to measured values for \emph{lac}, which should
be roughly the same because the gene function is similar. The transcription
rate for \emph{lac} in the absence of any repression that was found
in~\cite{RazoMejia2019} corresponds well with our estimate.

\paragraph*{The mRNA decay rate parameter $\gamma_{\n{mp}}$.}
The mRNA life time is usually on the order of 1 to 8 minutes which gives a
decay rate of $\gamma_{\n{m}} \approx 10^{-3} - 10^{-2}~\n{s^{-1}}$. Thus,
we have $\gamma_{\n{mp}} \defeq \frac{\gamma_{\n{m}}}{\gamma_{\n{p}}}
\approx 10^{1 \pm 0.5}$.

\paragraph*{The translation rate parameter $\rho_{\n{p}}$.}
The translation rate should be on the order of $r_{\n{p}} \approx 10^{-2} -
10^{-1}~\n{s^{-1}}$~\cite{Milo2016}. This gives $\rho_{\n{p}} \defeq
\frac{r_{\n{p}}}{\gamma_{\n{p}}} \approx 10^{2 \pm 0.5}$.

%\com{do growth rate thing}

\paragraph*{The Michaelis-Menten parameters for XapB ($k_{\n{\beta,i}}$, $K_{\n{\beta,i}}$, $k_{\n{\beta,e}}$, $K_{\n{\beta,e}}$):}
To estimate the parameters for both influx and efflux, we need data from
experiments where these two processes have been measured independently.
Lactose permease, which we expect to function in a similar way as XapB, is
comparatively well studied such that these rates have been measured in
multiple ways and for different conditions. Some data from several
publications is collected in~\cite{Viitanen1984} and we will base our
estimates on this and the corresponding references.

Comparing the effective rates and Michaelis constants of lactose permease
and the nucleoside transporters NupC and NupG shows that their orders of
magnitude are similar: from~\cite{Dornmair1989,Wright1985} we find
$k_{\n{cat}} \approx 10^1~\n{s^{-1}}$, $K_{\n{M}} \approx 10^1-10^2~\n{\mu
M}$ for LacY and from~\cite{Norholm2001,Komatsu1973,MunchPetersen1979}
$k_{\n{cat}} \approx 10^2~\n{s^{-1}}$, $K_{\n{M}} \approx 10^0-10^1~\n{\mu
M}$ for pyrimidines through NupC and NupG. Note that especially for Nup,
these are only rough measurements, and it is known that NupC behaves
different with purines (thus also with xanthosine). Given these values, we
expect lactose permease to be similar enough to nucleoside transporters so
that we can use it as a starting point for our estimation of the XapB
kinetic values.

We use the measured influx values that are summarized in~\cite{Viitanen1984}
for a membrane potential of $100~\n{mV}$ and obtain $k_{\n{b,i}} \approx
10^{1 \pm 1}~\n{s^{-1}}$ and $K_{\n{b,i}} \approx 10^{-4 \pm 2}~\n{M}$.
Here, the uncertainty is estimated from the distribution of values of many
kinds of enzymes that is presented in~\cite{Milo2016}. The bounds also
reflect the difference between LacY and Nup that was found above. For the
dimensionless quantities, this gives $k_{\n{\beta,i}} \approx 10^{4 \pm 1}$
and $K_{\n{\beta,i}} \approx 10^{1 \pm 2}$.

For the efflux parameters, we proceed similarly and obtain $k_{\n{b,e}}
\approx 10^{0 \pm 2}~\n{s^{-1}}$ and $K_{\n{b,e}} \approx 10^{-3 \pm
3}~\n{M}$, therefore $k_{\n{\beta,e}} \approx 10^{3 \pm 2}$ and
$K_{\n{\beta,e}} \approx 10^{2 \pm 3}$. Because of disagreement in
literature, difficulty in measuring these parameters, and a lack of
knowledge about what value range should be expected, we put larger error
bounds than for the influx.

\paragraph*{The non-specific transport parameters $k_{\n{\eta}}$, $\xi$:}
Estimating the rates of the Nup transporters for xanthosine is difficult
since there are no experimental values that could be used as a reference.
Because the rate cannot even be measured~\cite{Norholm2001}, it must be much
lower than that of XapB. The kinetic parameters should, however, be such
that the steady-state intracellular xanthosine concentration is
significantly larger than the extracellular one. Otherwise, there would be
no relevant difference to passive diffusion, which was shown to be
insufficient for switching~\cite{Norholm2001}. This essentially means $\xi <
1$, which on the other hand implies that the Nup rate is larger than or at
least equal to the diffusion rate: If Nup transport were slower than
diffusion, the latter would rule and $\xi$ would be close to~1 (because this
is the $\xi$ value of diffusion).

In addition, we know from \emph{lac} permease that influx is not much faster
than efflux, which implies that $\xi$ also cannot be too much smaller
than~1. We guess $\xi \approx 0.8 \pm 0.1$. The number of Nup transporters
per cell is roughly $10^2-10^3$~\cite{Li2014}. We expect
$\frac{k_{\n{cat}}}{K_{\n{M}}}$ to be at least two or three orders of
magnitude lower for Nup than for saturated XapB, where we have
$\frac{k_{\n{b}}}{K_{\n{b}}} \approx 10^{5 \pm 3}~\n{M^{-1}\ s^{-1}}$. These
arguments set an estimated upper bound on the non-specific transport rate:
$k_{\n{nup}} = \frac{k_{\n{cat}}}{K_{\n{M}}} \n{[E]_0} \lesssim 10^{-3 \pm
3}~\n{s^{-1}}$.

To find a lower bound, we estimate the diffusion rate. As described in
Eq~\ref{eq:diff} in the previous section, the diffusion rate can be written
as $10^{-2} \cdot p \cdot A$, where $p$ is the permeability coefficient of
the membrane for xanthosine, $A$ is the total cell surface area, and
$10^{-2}$ is the fraction of uncharged xanthosine at physiological pH. From
the permeabilities of various substances that are collected
in~\cite{Milo2016} and from the measured permeability values for purine
nucleosides in~\cite{Wei2011}, we estimate a permeability coefficient of
$\tilde{p}=10^{-8 \pm 3}~\n{m\ s^{{-1}}}$. We write the tilde because this
value is given in a convention where $\tilde{p} \cdot A = \dd{x}{t}$ with
$x$ being the number of xanthosine molecules and not the concentration.
Since we work in concentrations, we divide by an estimated cell volume of
$10^{-18}~\n{m^3}$ and, with a cell surface area of $10^{-11}~\n{m^2}$, find
a diffusion rate of roughly $10^{-3 \pm 3}~\n{s^{-1}}$. This sets the lower
bound of $k_{\n{nup}}$.
%\com{Check if Overton's rule gives the samediffusion coefficient}

Interestingly, the upper and lower bound turn out to be the same. Thus, we
obtain the range $k_{\n{nup}} \approx 10^{-3 \pm 3}~\n{s^{-1}}$, meaning
$k_{\n{\eta}} \approx 10^{-1 \pm 3}$.

\section{Additional plots and explanations of the results}
\paragraph*{The 2D set of equations.}
If the mRNA concentration is set to steady-state as discussed in the main
text, the dynamical system is reduced to the two equations shown below:
\begin{alignat}{2}
\dd{\mathrm{[p]_a}}{\tau} &= \ &&
\rho 
\frac{
	\mathrm{[XapR]_{R,A}}^2 
	e^{- \Delta \epsilon_{\text{coop}}}
}{
	1 + 
	2 \mathrm{[XapR]_{R,A}} +
	\mathrm{[XapR]_{R,A}}^2 e^{- \Delta \epsilon_{\text{coop}}}
}
- \mathrm{[p]_a}
\\
\dd{\mathrm{[x]_a}}{\tau} &= \ && \left(k_{\n{\beta,i}} \frac{\n{[c]_a}}{K_{\n{\beta,i}} + \n{[c]_a}} - k_{\n{\beta,e}} \frac{\n{[x]_a}}{K_{\n{\beta,e}} + \n{[x]_a}} - k_{\n{\alpha}} \frac{\n{[x]_a}}{1 + \n{[x]_a}}\right) \mathrm{[p]_a} \\ & && + k_{\n{\eta}} \left(\mathrm{[c]_a} - \xi \mathrm{[x]_a}\right)
\nonumber \\
\text{with} & && \raggedleft [\n{XapR}]_{\mathrm{R,\n{A}}} = [\n{XapR}]_{\mathrm{R}} \frac{\left(1 + \frac{\mathrm{[x]_{\n{a}}}}{K_{\n{\chi A}}}\right)^2}{\left(1 + \frac{\mathrm{[x]_{\n{a}}}}{K_{\n{\chi A}}}\right)^2 + e^{\Delta \epsilon_{\n{x}}} \left(1+\frac{\mathrm{[x]_{\n{a}}}}{K_{\n{\chi A}}} \frac{1}{K_{\n{IA}}}\right)^2}. \nonumber
\end{alignat}
Here, several of the parameters from the 3D system could now be combined
into one by defining $\rho \defeq \frac{\rho_{\n{m}}
\rho_{\n{p}}}{\gamma_{\n{mp}}}$.

\paragraph*{Magnitude of flow in the phase portraits.}
The plots in the main part do not show the magnitude of the vector fields.
For additional information, we show a 2D vector plot with the magnitudes in
Fig~\ref{figS4:vector}.

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{FigSI4.pdf}
	\caption{{\bf Vector plot of a standard case of bistability.}
		The 2D phase portraits as a vector plot for for the standard set of
		parameters that was used in the main text.}
	\label{figS4:vector} 
\end{figure}

\paragraph*{Increasing $\n{[c]_a}$.}
Fig~\ref{figS5:incC} shows a family of nullclines for increasing
extracellular xanthosine concentration. It shows the transition from
monostability, to bistability, and back to monostability again.

\begin{figure}
	\centering
	\includegraphics[width=0.8\textwidth]{FigSI5.pdf}
	\caption{{\bf A family of curves for increasing $\n{[c]_a}$.}
		The blue curve is the protein nullcline, which does not change with
		$\n{[c]_a}$. The other lines are the xanthosine nullclines for
		different values of the extracellular xanthosine concentration.}
	\label{figS5:incC}
\end{figure}

\paragraph*{Hysteresis effects.}
In Fig~\ref{figS8:hysteresis}, the results mentioned in the hysteresis part
of the main text are shown. As mentioned there,  the simulation was started
with initially fully induced cells. At concentrations where no cells
switched from the low to the high expression state before, cells here do not
switch back to low expression.

\begin{figure}
	\centering
	\includegraphics[width=0.9\textwidth]{FigSI8.pdf}
	\caption{{\bf Decreasing $\n{[c]_a}$ in the fully induced system.}
		Apart from $\n{[c]_a}$, the parameters are the same as in the table
		in the main text. For the distributions, the simulations were run
		1000 times for $10^6 \unit{s}$ each and started at the mRNA, protein
		and intracellular xanthosine counts of the high fixed point in the
		corresponding phase portraits. The values of $\n{[c]_a}$ are 12 in
		(A), 10 in (B), and 9 in (C). When comparing to the distributions in
		the main text where the simulation had initial concentrations of 0,
		a clear hysteresis effect can be observed.}
	\label{figS8:hysteresis}
\end{figure}

\section{Chemical master equation}
Here we provide some further details on our stochastic simulations by
writing the chemical master equation for our model. The state of our system
is described by the number of mRNA $m$, protein $p$, and xanthosine
molecules $x$, and we write $P(m,p,x)$ for the probability of that state.
The chemical master equation for the system is:
\begin{alignat}{2}
\dd{P(m,p,x)}{t} &= \ && - P(m,p,x) [ r_{\n{m}} p_{\n{active}}(x) + \gamma_{\n{m}} m + r_{\n{p}} m + \gamma_{\n{p}} p \nonumber \\ & && + k_{\n{b,i}} \frac{c}{K_{\n{b,i}} + c} \n{p} + k_{\n{b,e}} \frac{x}{K_{\n{b,e}} + x} p + k_{\n{a}} \frac{x}{K_{\n{a}} + x} p \nonumber \\ & && + k_{\n{nup}} c + k_{\n{nup}} \xi x ] \nonumber \\ & && 
+ P(m+1,p,x) \gamma_{\n{m}} (m + 1) \nonumber \\ & &&  + P(m-1,p,x) r_{\n{m}} p_{\n{active}}(x) \nonumber \\ & && 
+ P(m,p+1,x) \gamma_{\n{p}} (p + 1) \nonumber \\ & &&  + P(m,p-1,x) r_{\n{p}} m \nonumber \\ & && 
+ P(m,p,x+1) [ k_{\n{b,e}} \frac{x+1}{K_{\n{b,e}} + x+1} p \nonumber \\ & && + k_{\n{a}} \frac{x+1}{K_{\n{a}} + x+1} p + k_{\n{nup}} \xi (x+1) ] \nonumber \\ & &&
+ P(m,p,x-1) [ k_{\n{b,i}} \frac{c}{K_{\n{b,i}} + c} p + k_{\n{nup}} c ]  
\end{alignat}
Here, all the parameters of course have to be measured in counts, not in
concentrations. As can be seen from the equation, none of the propensities
depends explicitly on time and hence, we can use the hybrid algorithm
between classical Gillespie and \texttau-leaping that we relied on.

\FloatBarrier

\begin{thebibliography}{10}

	\bibitem{Phillips2018}
	{Phillips} R, {Belliveau} NM, {Chure} G, {Garcia} HG, {Razo-Mejia} M, {Scholes}
	C.
	\newblock {Figure 1 Theory Meets Figure 2 Experiments in the Study of Gene
		Expression}.
	\newblock arXiv e-prints. 2018; p. arXiv:1812.11627.

	\bibitem{RazoMejia2018}
	Razo-Mejia M, Barnes SL, Belliveau NM, Chure G, Einav T, Lewis M, et~al.
	\newblock Tuning Transcriptional Regulation through Signaling: A Predictive
	Theory of Allosteric Induction.
	\newblock Cell Systems. 2018;6(4):456 -- 469.e10.
	\newblock doi:{10.1016/j.cels.2018.02.004}.

	\bibitem{Chure2019}
	Chure G, et~al.
	\newblock Measurements on the \emph{xapABR} system in \emph{E. coli.};
	Unpublished.

	\bibitem{Norholm2001}
	Norholm MH, Dandanell G.
	\newblock Specificity and topology of the \emph{Escherichia coli} xanthosine
	permease, a representative of the NHS subfamily of the major facilitator
	superfamily.
	\newblock Journal of bacteriology. 2001;183(11466294):4900--4904.

	\bibitem{Kaback2015}
	Kaback HR.
	\newblock A chemiosmotic mechanism of symport.
	\newblock Proceedings of the National Academy of Sciences.
	2015;112(5):1259--1264.
	\newblock doi:{10.1073/pnas.1419325112}.

	\bibitem{Kaczorowski1979}
	Kaczorowski GJ, Robertson DE, Kaback HR.
	\newblock Mechanism of lactose translocation in membrane vesicles from
	\emph{Escherichia coli}. 2. Effect of imposed $\Delta \Psi$, $\Delta$ pH, and
	$\Delta \mu_{\mathrm{H^+}}$.
	\newblock Biochemistry. 1979;18(17):3697--3704.
	\newblock doi:{10.1021/bi00584a010}.

	\bibitem{Viitanen1984}
	Viitanen P, Garcia ML, Kaback HR.
	\newblock Purified reconstituted lac carrier protein from \emph{Escherichia
		coli} is fully functional.
	\newblock Proceedings of the National Academy of Sciences of the United States
	of America. 1984;81(6324209):1629--1633.

	\bibitem{Dandanell2005}
	Dandanell G, Szczepanowski RH, Kierdaszuk B, Shugar D, Bochtler M.
	\newblock \emph{Escherichia coli} Purine Nucleoside Phosphorylase II, the
	Product of the xapA Gene.
	\newblock Journal of Molecular Biology. 2005;348(1):113--125.

	\bibitem{Koszalka1988}
	Koszalka GW, Vanhooke J, Short SA, Hall WW.
	\newblock Purification and properties of inosine-guanosine phosphorylase from
	\emph{Escherichia coli} K-12.
	\newblock Journal of bacteriology. 1988;170(3042752):3493--3498.

	\bibitem{Li2014}
	Li GW, Burkhardt D, Gross C, Weissman J.
	\newblock Quantifying Absolute Protein Synthesis Rates Reveals Principles
	Underlying Allocation of Cellular Resources.
	\newblock Cell. 2014;157(3):624 -- 635.
	\newblock doi:{https://doi.org/10.1016/j.cell.2014.02.033}.

	\bibitem{Milo2016}
	Milo R, Phillips R.
	\newblock In: Cell biology by the numbers. Garland Science; 2016. p. 217.

	\bibitem{RazoMejia2019}
	Razo-Mejia M, Phillips R.
	\newblock First-principles prediction of the information processing capacity of
	a simple genetic circuit.
	\newblock bioRxiv. 2019;doi:{10.1101/594325}.

	\bibitem{Dornmair1989}
	Dornmair K, Overath P, Jähnig F.
	\newblock Fast measurement of galactoside transport by lactose permease.
	\newblock The Journal of biological chemistry. 1989;264:342--6.

	\bibitem{Wright1985}
	Wright J K, Seckler R.
	\newblock The lactose/H+ carrier of \emph{Escherichia coli}: lac YUN mutation
	decreases the rate of active transport and mimics an energy-uncoupled
	phenotype.
	\newblock The Biochemical journal. 1985;227:287--97.
	\newblock doi:{10.1042/bj2270287}.

	\bibitem{Komatsu1973}
	Komatsu Y, Tanaka K.
	\newblock Deoxycytidine uptake by isolated membrain vesicles from
	\emph{Escherichia coli} K 12.
	\newblock Biochimica et Biophysica Acta (BBA) - Biomembranes.
	1973;311(4):496--506.

	\bibitem{MunchPetersen1979}
	Munch-Petersen A, Mygind B, Nicolaisen AK, Pihl NJ.
	\newblock Nucleoside transport in cells and membrane vesicles from
	\emph{Escherichia coli} K12.
	\newblock The Journal of biological chemistry. 1979;254 10:3730--7.

	\bibitem{Wei2011}
	Wei C, Pohorille A.
	\newblock Permeation of Nucleosides through Lipid Bilayers.
	\newblock J Phys Chem B. 2011;115(13):3681--3688.
	\newblock doi:{10.1021/jp112104r}.

\end{thebibliography}

\end{document}